---
title: "p8105_hw2_pl2811"
author: "Pei Hsin Lin"
date: "10/9/2021"
output: html_document
---
```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggridges)
library(lubridate)
```

```{r}
###Problem 1
#Read and clean the Mr. Trash Wheel sheet
trash_wheel<-read_excel("/Users/lin/Desktop/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, col_names = TRUE, col_types = NULL, na = "", skip = 0)
names(trash_wheel)
trash_wheel= janitor::clean_names(trash_wheel)
names(trash_wheel)
skimr::skim(trash_wheel)
trash_wheel= subset(trash_wheel, select = -c(x15,x16,x17) )
trash_wheel<- na.omit(trash_wheel)

#round sports balls to the nearest integer
  trash_wheel %>%
pull(sports_balls)%>%
round()

#Read and clean precipitation data for 2018 
precipitation_2018<-read_excel("/Users/lin/Desktop/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", col_names = TRUE, col_types = NULL, na = "", skip = 1)
precipitation_2018<- na.omit(precipitation_2018)
#add a variable for year. 
precipitation_2018<- precipitation_2018 %>%
  add_column(year = "2018")
#combine datasets,  convert month to a character variable 
month<-month.name[precipitation_2018$Month]
precipitation_2018<- precipitation_2018 %>%
add_column(month)
precipitation_2018= subset(precipitation_2018, select = -c(Month))
precipitation_2018<- precipitation_2018[c("year", "month", "Total")]
precipitation_2018 %>%
pull(Total)%>%
sum()
#Read and clean precipitation data for 2019
precipitation_2019<-read_excel("/Users/lin/Desktop/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", col_names = TRUE, col_types = NULL, na = "", skip = 1)
precipitation_2019<- na.omit(precipitation_2019)
#add a variable for year.
precipitation_2019<- precipitation_2019 %>%
  add_column(year = "2019")
#combine datasets,  convert month to a character variable  
month<-month.name[precipitation_2019$Month]
precipitation_2019<- precipitation_2019 %>%
add_column(month)
precipitation_2019= subset(precipitation_2019, select = -c(Month))
precipitation_2019<- precipitation_2019[c("year", "month", "Total")]
summary(precipitation_2019)
balls_19=subset(trash_wheel, select = c(sports_balls, year))

balls_19<-filter(balls_19,year == 2019)
summary(balls_19)

```

Mr. Trash Wheel sheet has 453 observations, precipitation data for 2018 has 12 observations,precipitation data for 2019 has 12 observations.
Examples of key variables include weight_tons, volume_cubic_yards, and plastic_bottles etc.
Total precipitation in 2018 is 70.33, median number of sports balls in a dumpster in 2019 is 9.


###Problem 2

```{r}

#clean the data, break up the variable,
pols_month<- read.csv(file ="/Users/lin/Desktop/fivethirtyeight_datasets/pols-month.csv")
pols_month<- pols_month %>%
    separate(mon, c("year", "Month", "day"), sep="-")
#replace month number with month name
pols_month$Month <- as.numeric(as.character(pols_month$Month ))
month<-month.name[pols_month$Month]
pols_month<- pols_month %>%
add_column(month)
pols_month= subset(pols_month, select = -c(Month))
###create variable, remove the day variable.
pols_month= 
  pivot_longer(
    pols_month, 
    starts_with("prez"),
    names_to = "president", 
    names_prefix = "prez_",
    values_to = "prez")
pols_month= subset(pols_month, select = -c(day))

```

```{R}
###clean the data using a similar process to the above
snp<- read.csv(file ="/Users/lin/Desktop/fivethirtyeight_datasets/snp.csv")
snp_y<- mdy(snp$date)
foo <- function(x, year=1968){
  m <- year(x) %% 100
  year(x) <- ifelse(m > year %% 100, 1900+m, 2000+m)
  x
}
Date<-foo(snp_y, 1949)
snp<- snp %>%
add_column(Date)
snp= subset(snp, select = -c(date))
snp<- snp %>%
    separate(Date, c("year", "Month", "day"), sep="-")
snp= subset(snp, select = -c(day))
snp$Month <- as.numeric(as.character(snp$Month ))
month<-month.name[snp$Month]
snp<- snp %>%
add_column(month)
snp= subset(snp, select = -c(Month))
###organize so that year and month are the leading columns
snp<- snp[c("year", "month", "close")]
```

```{r}
###tidy the unemployment data, merge datasets
unemployment<-read.csv(file ="/Users/lin/Desktop/fivethirtyeight_datasets/unemployment.csv")
unemployment <- unemployment%>% rename( January = Jan, February=Feb, March=Mar,April=Apr, June=Jun, July=Jul,  August=Aug, September=Sep, October=Oct, November=Nov,December=Dec)
unemployment<- gather(unemployment, month, unemployment,January:December, factor_key=TRUE)
unemployment
unemployment<-janitor::clean_names(unemployment)
snp<-janitor::clean_names(snp)
pols_month<-janitor::clean_names(pols_month)
###merging snp into pols, and merging unemployment
total <- merge(pols_month,snp,by=c("year","month"))
total <- merge(total,unemployment, by=c("year","month"))
```
The total dataset is a combined of unemployment, snp, pols_month data sets, 
including variables year,month,gov_gop,sen_gop,rep_gop,gov_dem,sen_dem,rep_dem,president,prez, close. total data sets has 1572 rows, 12 columns. Range of year from 1950 to 2015.


```{r}
###Problem 3

#Load and tidy the data
Popular_Baby_Names<-read.csv(file ="/Users/lin/Desktop/Popular_Baby_Names.csv")
Popular_Baby_Names<-Popular_Baby_Names[!duplicated(Popular_Baby_Names), ]
Popular_Baby_Names<-janitor::clean_names(Popular_Baby_Names)

###Produce table showing the rank in popularity of the name “Olivia” 
Olivia<-filter(Popular_Baby_Names, child_s_first_name == "Olivia")
Olivia= subset(Olivia, select = -c(gender,child_s_first_name,count))
Olivia<-Olivia %>%
  pivot_wider( names_from = ethnicity, 
  values_from = rank)
Olivia<-arrange(Olivia,year_of_birth)
Olivia<-janitor::clean_names(Olivia)
Olivia
#Producetable showing the most popular name among male children over time.
p_male<-filter(Popular_Baby_Names,gender == "MALE")
p_male<-filter(p_male,rank ==1)
p_male=subset(p_male, select = -c(gender,count,rank))
p_male<-p_male%>%
  pivot_wider( names_from = ethnicity, 
  values_from =child_s_first_name)
p_male<-janitor::clean_names(p_male)
p_male$asian_and_pacific_islander<- coalesce( p_male$asian_and_pacific_islander,  p_male$asian_and_paci)
p_male$ black_non_hispanic<- coalesce( p_male$ black_non_hispanic,  p_male$ black_non_hisp)
p_male$white_non_hispanic<- coalesce( p_male$white_non_hispanic,  p_male$white_non_hisp)
p_male=subset(p_male, select = c(year_of_birth ,asian_and_pacific_islander,black_non_hispanic,white_non_hispanic,hispanic))
p_male<-arrange(p_male, year_of_birth)
p_male
###f produce a scatter plot showing the number of children with a name against the rank in popularity of that name

mw16<-filter(Popular_Baby_Names,gender == "MALE" &  year_of_birth == 2016)
mw16<-filter(mw16,ethnicity=="WHITE NON HISPANIC"|ethnicity=="WHITE NON HISP")

              
mw16_plot<-ggplot(mw16, aes(x = rank, y = count)) + 
    geom_point()
mw16_plot

```


